---
output: github_document
---

<!-- README.md is generated from README.Rmd/ Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse=TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  dev = 'svg',
  out.width = "100%")
options(tibble.print_min = 5, tibble.print_max = 5)
```

# denoisr

<!-- badges: start -->
[![Build Status](https://travis-ci.org/StevenGolovkine/denoisr.svg?branch=master)](https://travis-ci.org/StevenGolovkine/denoisr)
[![Codacy Badge](https://api.codacy.com/project/badge/Grade/d5b2b6c6083c4d269ace4a8a1b8103b3)](https://www.codacy.com/manual/StevenGolovkine/denoisr?utm_source=github.com&amp;utm_medium=referral&amp;utm_content=StevenGolovkine/denoisr&amp;utm_campaign=Badge_Grade)
<!-- badges: end -->

## Overview

`denoisr` is a non-parametric smoother for noisy curve data, providing different
functions to estimate various parameters:

* `estimate_H0_list()` and `estimate_H0_deriv_list()` estimate the smoothness of
the curves.
* `estimate_b_list()` and `estimate_bandwidth()` estimate the bandwidth used in
the Nadaraya-Watson estimator.
* `estimate_curve()` estimates one curve, given bandwidths.
* `smooth_curves()` and `smooth_curves_regularity()` estimate the curves.
* `estimate_mean()` and `estimate_covariance()` estimate the leave-one-out
smoothing mean and covariance.

You can learn more about thm in `vignette('denoisr')`.

## Installation

The package is not yet on CRAN.

## Development version

You can install the last released version of `denoisr` from [Github](https://github.com/StevenGolovkine/denoisr) with:

```{r, eval=FALSE}
devtools::install_github("StevenGolovkine/denoisr")
```

## Usage

```{r load_package, warning=FALSE, message=FALSE}
# Load packages
library(denoisr)
library(dplyr)
library(funData)
library(ggplot2)
library(reshape2)
```

```{r load_data}
# Load data
data("canadian_temperature_daily")
```

```{r convert_funData}
# Convert list to funData object
data_fd <- list2funData(canadian_temperature_daily)
```

```{r plot_data, fig.height=4.35, fig.width=7, fig.align='center'}
# Plot of the data
autoplot(data_fd) + 
  labs(title = 'Daily Temperature Data',
       x = 'Normalized Day of Year',
       y = 'Temperature in 째C') +
  theme_minimal()
```

```{r smooth_data}
# Smooth the data
smooth_data <- smooth_curves(canadian_temperature_daily, 
                             t0_list = 0.5, 
                             k0_list = 5)
```

```{r plot_data2, fig.height=4.35, fig.width=7, fig.align='center'}
# Plot of the smoothed data
smooth_data_fd <- list2funData(smooth_data$smooth)
autoplot(smooth_data_fd) +
  labs(title = 'Smooth Daily Temperature Data',
       x = 'Normalized Day of Year',
       y = 'Temperature in 째C') +
  theme_minimal()
```

```{r plot_data3, fig.height=4.35, fig.width=7, fig.align='center'}
# Plot one realisation of the smoothed data
montreal <- tibble(t = canadian_temperature_daily$Montreal$t,
                   Data = canadian_temperature_daily$Montreal$x,
                   Smooth = smooth_data$smooth$Montreal$x) %>% 
  reshape2::melt(id = 't')
ggplot(montreal, aes(x = t, y = value, color = variable)) +
  geom_line() +
  labs(title = 'Example of Montreal',
       x = 'Normalized Day of Year',
       y = 'Temperature in 째C',
       color = '') +
  theme_minimal()

```

```{r compute_mean}
# Compute the mean curve
mean_curve <- estimate_mean(canadian_temperature_daily,
                            U = seq(0, 1, length.out = 501),
                            b = smooth_data$parameter$b)
```

```{r plot_mean, fig.height=4.35, fig.width=7, fig.align='center'}
mean_curve_plot <- tibble(t = seq(0, 1, length.out = 501), x = mean_curve)
autoplot(data_fd) +
  geom_line(data = mean_curve_plot, mapping = aes(x = t, y = x, group = NULL), col = 'red', size = 2) +
  labs(title = 'Daily Temperature Data',
       x = 'Normalized Day of Year',
       y = 'Temperature in 째C') +
  theme_minimal()
```

```{r compute_covariance}
# Compute the covariance surface
U <- seq(0, 1, length.out = 101)
cov_surface <- estimate_covariance(canadian_temperature_daily,
                                   U = U,
                                   b = smooth_data$parameter$b,
                                   h = 0.05)
```

```{r plot_covariance, fig.height=5, fig.width=5, fig.align='center'}
cov_surface_plot <- tibble(expand.grid(U, U), C = as.vector(cov_surface))
ggplot(cov_surface_plot) + 
  geom_contour_filled(aes(x = Var1, y = Var2, z = C)) + 
  labs(title = 'Covariance surface of daily temperature data',
       x = '',
       y = '',
       fill = 'Level') +
  theme_minimal()
```