---
title: "Introduction to denoisr"
author: "Steven Golovkine"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to denoisr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*References:*

  * J. O. Ramsay, Spencer Graves and Giles Hooker (2020). fda: Functional Data Analysis. R package version 5.1.4. https://CRAN.R-project.org/package=fda
  * Happ-Kurz C (2020). “Object-Oriented Software for Functional Data.” _Journal of Statistical Software_, *93*(5), 1-38. doi: 10.18637/jss.v093.i05 (URL: https://doi.org/10.18637/jss.v093.i05).

## Introduction

`denoisr` is a **R** package which permits to smooth (remove the noise from) functional
data by, first, estimate the Hurst coefficient of the underlying generating
process. 

Functional data to smooth should be defined on a univariate compact, but can be
irregularly sampled. `denoisr` can also be used only for Hurst parameter 
estimation. And finally, there is also some functions to compute the mean curve and the corariance surface.

`denoisr` have a support for the package [`funData`](https://cran.r-project.org/web/packages/funData/index.html).

## Model

Let $I$ be a compact interval of $\mathbb{R}$. We assume $X^{(1)}, \dots, X^{(N)}$ to be an independent sample of a random process $X = \left(X_t : t \in I\right)$ with continuous trajectories. For each $1 \leq n \leq N$, and given a positive integer $M_n$, let $T_m^{(n)}, 1 \leq m \leq M_n$ be the random observation times for the curve $X^{(n)}$. These times are obtained as independent copies of a variable $T$ taking values in $I$. The integers $M_1, \dots, M_N$ represent an independent sample of an integer-valued random variable $M$ with expectation $\mu$. We assume that the realizations $X$, $M$ and $T$ are mutually independent. The observations consist of the pairs $\left(Y_m^{(n)}, T_m^{(n)}\right) \in \mathbb{R} \times I$ where $Y_m^{(n)}$ is defined as
$$Y_m^{(n)} = X^{(n)}(T_m^{(n)}) + \varepsilon_m^{(n)}, \quad 1 \leq n \leq N,~ 1 \leq m \leq M_n,$$
and $\varepsilon_m^{(n)}$ are independent copies of a centered error variable $\varepsilon$.

 Let $t_0 \in I$ an arbitrarily fixed piont. The aim is to estimate $X^{(1)}(t_0), \dots, X^{(N)}(t_0)$.

```{r setup, warning=FALSE, message=FALSE}
# Load packages
library(denoisr)
library(dplyr)
library(funData)
library(ggplot2)
library(reshape2)
```

## Example on temperature data 

The Canadian temperature data correspond to the daily temperature at $35$ different locations in Canada averaged over 1960 to 1994. This dataset is included in the package [`fda`](https://CRAN.R-project.org/package=fda).

This section performs the smoothing of the Canadian temperature Weather data.


```{r load_data}
# Load data
data("canadian_temperature_daily")
```

This package manipulates lists to define functional data. So, functions are provided to convert it as functional data object defined in the package *funData*.

```{r convert_funData}
# Convert list to funData object
data_fd <- list2funData(canadian_temperature_daily)
```

So, we can use all the available function for `funData` object. As an example, we can plot the data with the `ggplot2` style.

```{r plot_data, fig.height=4.35, fig.width=7}
# Plot of the data
autoplot(data_fd) + 
  labs(title = 'Daily Temperature Data',
       x = 'Normalized Day of Year',
       y = 'Temperature in °C') +
  theme_minimal()
```


The smoothing of the curves can be performed using the function `smooth_curves`. It accepts the parameter `t0_list` which correspond to the times where we want to estimate the bandwidth and the parameter `k0_list` which is the considered neighborhood.

```{r smooth_data}
# Smooth the data
smooth_data <- smooth_curves(canadian_temperature_daily, 
                             t0_list = 0.5, 
                             k0_list = 5)
```

Then, we can plot the smoothed data.
```{r plot_data2, fig.height=4.35, fig.width=7}
# Plot of the smoothed data
smooth_data_fd <- list2funData(smooth_data$smooth)
autoplot(smooth_data_fd) +
  labs(title = 'Smooth Daily Temperature Data',
       x = 'Normalized Day of Year',
       y = 'Temperature in °C') +
  theme_minimal()
```

Finally, we show on particular curve along its smoothed version to look at the impact of smoothing.
```{r plot_data3, fig.height=4.35, fig.width=7}
# Plot one realisation of the smoothed data
montreal <- tibble(t = canadian_temperature_daily$Montreal$t,
                   Data = canadian_temperature_daily$Montreal$x,
                   Smooth = smooth_data$smooth$Montreal$x) %>% 
  reshape2::melt(id = 't')
ggplot(montreal, aes(x = t, y = value, color = variable)) +
  geom_line() +
  labs(title = 'Example of Montreal',
       x = 'Normalized Day of Year',
       y = 'Temperature in °C',
       color = '') +
  theme_minimal()

```

### Mean and Covariance

First, we compute the mean curve. The mean curve is defined as the mean of the Leave-One-Out mean curves that are computed after removing one curve sequentially.
```{r compute_mean}
# Compute the mean curve
mean_curve <- estimate_mean(canadian_temperature_daily,
                            U = seq(0, 1, length.out = 501),
                            b = smooth_data$parameter$b)
```

Show the mean function along the complete set of curves.
```{r plot_mean, fig.height=4.35, fig.width=7}
mean_curve_plot <- tibble(t = seq(0, 1, length.out = 501), x = mean_curve)
autoplot(data_fd) +
  geom_line(data = mean_curve_plot, mapping = aes(x = t, y = x, group = NULL), col = 'red', size = 2) +
  labs(title = 'Daily Temperature Data',
       x = 'Normalized Day of Year',
       y = 'Temperature in °C') +
  theme_minimal()
```

Next, we compute the covariance surface. The covariance surface is defined as a 2-dimensional Nadaraya-Watson estimator of the smoothed version of the noisy curves.
```{r compute_covariance}
# Compute the covariance surface
U <- seq(0, 1, length.out = 101)
cov_surface <- estimate_covariance(canadian_temperature_daily,
                                   U = U,
                                   b = smooth_data$parameter$b,
                                   h = 0.05)
```

Show the covariance surface.
```{r plot_covariance, fig.height=5, fig.width=5, fig.align='center'}
cov_surface_plot <- tibble(expand.grid(U, U), C = as.vector(cov_surface))
ggplot(cov_surface_plot) + 
  geom_contour_filled(aes(x = Var1, y = Var2, z = C)) + 
  labs(title = 'Covariance surface of daily temperature data',
       x = '',
       y = '',
       fill = 'Level') +
  theme_minimal()
```



## Example on simulated data

We will do the same analysis on some simulated datasets.


### On fractional brownian motion

First, we will do the analysis on a dataset of fractional brownian motion. A fractional brownian motion is a continuous-time Gaussian process $B_H(t)$ on $[0, 1]$, such that:

  * $B_H(0) = 0$

  * $\mathbb{E}(B_H(t)) = 0, \quad \forall t \in [0, 1]$

  * $\mathbb{E}(B_H(s)B_H(t)) = \frac{1}{2}\left(\lvert s\rvert^{2H} + \lvert t\rvert^{2H} - \lvert t - s\rvert^{2H}\right)$ where $H$ is a real number in $(0, 1)$, named Hurst parameter.
  

```{r load_brownian, warning=FALSE, message=FALSE}
# Simulate some data
set.seed(42)
fractional_brownian <- generate_fractional_brownian(N = 100, M = 350, 
                                                    H = 0.5, sigma = 0.05)
```

```{r smooth_data_brown}
# Smooth the data
smooth_data <- smooth_curves(fractional_brownian, 
                             t0_list = 0.5, 
                             k0_list = 14)
```

```{r estimate_H_brow}
# Estimation of the Hurst coefficient
print(smooth_data$parameter$H0)
```


```{r plot_brown, fig.height=4.35, fig.width=7}
# Plot a particular observation
obs <- tibble(t = fractional_brownian[[1]]$t,
              Noisy = fractional_brownian[[1]]$x,
              Truth = fractional_brownian[[1]]$x_true,
              Smooth = smooth_data$smooth[[1]]$x) %>% 
  reshape2::melt(id = 't')
ggplot(obs, aes(x = t, y = value, color = variable)) +
  geom_line() +
  labs(title = 'Fractional brownian motion',
       x = 'Normalized time',
       y = 'Value',
       color = '') +
  theme_minimal()
```

```{r mean_frac_brown}
# Compute the mean curve
mean_curve <- estimate_mean(fractional_brownian,
                            U = seq(0, 1, length.out = 100),
                            b = smooth_data$parameter$b)
```

```{r plot_mean_frac_brown, fig.height=4.35, fig.width=7}
data_fd <- list2irregFunData(fractional_brownian)
mean_curve_plot <- tibble(t = seq(0, 1, length.out = 100), x = mean_curve)
autoplot(data_fd) +
  geom_line(data = mean_curve_plot, mapping = aes(x = t, y = x, group = NULL), col = 'red', size = 2) +
  labs(title = 'Fractional brownian motion',
       x = 'Normalized time',
       y = 'Value') +
  theme_minimal()
```

```{r compute_covariance_frac_brown}
# Compute the covariance surface
U <- seq(0, 1, length.out = 101)
cov_surface <- estimate_covariance(fractional_brownian,
                                   U = U,
                                   b = smooth_data$parameter$b,
                                   h = 0.05)
```

```{r plot_covariance_frac_brown, fig.height=5, fig.width=5, fig.align='center'}
# Show the covariance surface
cov_surface_plot <- tibble(expand.grid(U, U), C = as.vector(cov_surface))
ggplot(cov_surface_plot) + 
  geom_contour_filled(aes(x = Var1, y = Var2, z = C)) + 
  labs(title = 'Covariance surface of fractional brownian motion',
       x = '',
       y = '',
       fill = 'Level') +
  theme_minimal()
```


### On piecewise fractional brownian motion

```{r load_piecewise_brownian, warning=FALSE, message=FALSE}
set.seed(42)
piece_frac_brown <- generate_piecewise_fractional_brownian(N = 150, M = 350, 
                                                           H = c(0.2, 0.5, 0.8), 
                                                           sigma = 0.05)
```

```{r smooth_data_piecebrown}
# Smooth the data
smooth_data <- smooth_curves(piece_frac_brown, 
                             t0_list = c(0.15, 0.5, 0.85), 
                             k0_list = c(14, 14, 14))
```

```{r estimate H_piecebrow}
# Estimation of the Hurst coefficient
print(smooth_data$parameter$H0)
```

```{r plot_piece_brown, fig.height=4.35, fig.width=7}
# Plot a particular observation
obs <- tibble(t = piece_frac_brown[[1]]$t,
              Noisy = piece_frac_brown[[1]]$x,
              Truth = piece_frac_brown[[1]]$x_true,
              Smooth = smooth_data$smooth[[1]]$x) %>% 
  reshape2::melt(id = 't')
ggplot(obs, aes(x = t, y = value, color = variable)) +
  geom_line() +
  labs(title = 'Piecewise fractional brownian motion',
       x = 'Normalized time',
       y = 'Value',
       color = '') +
  theme_minimal()
```

```{r mean_piece_frac_brown}
# Compute the mean curve
mean_curve <- estimate_mean(piece_frac_brown,
                            U = seq(0, 1, length.out = 100),
                            b = smooth_data$parameter$b,
                            t0_list = c(0.15, 0.5, 0.85))
```

```{r plot_mean_piece_frac_brown, fig.height=4.35, fig.width=7}
data_fd <- list2irregFunData(piece_frac_brown)
mean_curve_plot <- tibble(t = seq(0, 1, length.out = 100), x = mean_curve)
autoplot(data_fd) +
  geom_line(data = mean_curve_plot, mapping = aes(x = t, y = x, group = NULL), col = 'red', size = 2) +
  labs(title = 'Fractional brownian motion',
       x = 'Normalized time',
       y = 'Value') +
  theme_minimal()
```

```{r compute_covariance_piece_frac_brown}
# Compute the covariance surface
U <- seq(0, 1, length.out = 101)
cov_surface <- estimate_covariance(piece_frac_brown,
                                   U = U,
                                   b = smooth_data$parameter$b,
                                   h = 0.05,
                                   t0_list = c(0.15, 0.5, 0.85))
```

```{r plot_covariance_piece_frac_brown, fig.height=5, fig.width=5, fig.align='center'}
# Show the covariance surface
cov_surface_plot <- tibble(expand.grid(U, U), C = as.vector(cov_surface))
ggplot(cov_surface_plot) + 
  geom_contour_filled(aes(x = Var1, y = Var2, z = C)) + 
  labs(title = 'Covariance surface of \n piecewise fractional brownian motion',
       x = '',
       y = '',
       fill = 'Level') +
  theme_minimal()
```

### On integrated fractional brownian motion

```{r load_inte_brownian, warning=FALSE, message=FALSE}
# Simulate some data
set.seed(42)
inte_fractional_brownian <- generate_integrate_fractional_brownian(N = 100, M = 350, H = 0.5, sigma = 0.025)
```

```{r smooth_data_inte_brown}
# Smooth the data
smooth_data <- smooth_curves_regularity(inte_fractional_brownian, 
                                        t0 = 0.5, k0 = 14)
```

```{r estimate_H_inte_brow}
# Estimation of the Hurst coefficient
print(smooth_data$parameter$H0)
```


```{r plot_inte_brown, fig.height=4.35, fig.width=7}
# Plot a particular observation
obs <- tibble(t = inte_fractional_brownian[[1]]$t,
              Noisy = inte_fractional_brownian[[1]]$x,
              Truth = inte_fractional_brownian[[1]]$x_true,
              Smooth = smooth_data$smooth[[1]]$x) %>% 
  reshape2::melt(id = 't')
ggplot(obs, aes(x = t, y = value, color = variable)) +
  geom_line() +
  labs(title = 'Integreted Fractional brownian motion',
       x = 'Normalized time',
       y = 'Value',
       color = '') +
  theme_minimal()
```

```{r mean_frac_inte_brown}
# Compute the mean curve
mean_curve <- estimate_mean(inte_fractional_brownian,
                            U = seq(0, 1, length.out = 100),
                            b = smooth_data$parameter$b)
```

```{r plot_mean_inte_frac_brown, fig.height=4.35, fig.width=7}
data_fd <- list2irregFunData(inte_fractional_brownian)
mean_curve_plot <- tibble(t = seq(0, 1, length.out = 100), x = mean_curve)
autoplot(data_fd) +
  geom_line(data = mean_curve_plot, mapping = aes(x = t, y = x, group = NULL), col = 'red', size = 2) +
  labs(title = 'Integrated fractional brownian motion',
       x = 'Normalized time',
       y = 'Value') +
  theme_minimal()
```

```{r compute_covariance_inte_frac_brown}
# Compute the covariance surface
U <- seq(0, 1, length.out = 101)
cov_surface <- estimate_covariance(inte_fractional_brownian,
                                   U = U,
                                   b = smooth_data$parameter$b,
                                   h = 0.05)
```

```{r plot_covariance_inte_frac_brown, fig.height=5, fig.width=5, fig.align='center'}
# Show the covariance surface
cov_surface_plot <- tibble(expand.grid(U, U), C = as.vector(cov_surface))
ggplot(cov_surface_plot) + 
  geom_contour_filled(aes(x = Var1, y = Var2, z = C)) + 
  labs(title = 'Covariance surface of integrated fractional brownian motion',
       x = '',
       y = '',
       fill = 'Level') +
  theme_minimal()
```

