---
title: "Introduction to `denoisr`"
author: "Steven Golovkine"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{denoisr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*References:*

  * funData: An S4 Class for Functional Data, Clara Happ-Kurz, 2019, https://CRAN.R-project.org/package=funData

# Introduction

`denoisr` is a *R* package which permits to smooth (remove the noise) functional
data by, first, estimate the Hurst coefficient of the underlying generating
process. 

Functional data to smooth should be defined on a univariate compact, but can be
irregularly sampled. `denoisr` can also be used only for Hurst parameter 
estimation.

`denoisr` have a support for the package [`funData`](https://cran.r-project.org/web/packages/funData/index.html).

## Example on temperature data 

This section performs the smoothing of the Canadian temperature Weather data.

```{r setup, warning=FALSE, message=FALSE}
# Load packages
library(denoisr)
library(funData)
```


```{r load_data}
data("canadian_temperature_daily")
```

```{r convert_funData}
data_fd <- list2funData(canadian_temperature_daily)
```


```{r plot_data}
plot(data_fd)
```

```{r smooth_data}
smooth_data <- smooth_curves(canadian_temperature_daily, 
                             t0_list = 0.5, 
                             k0_list = 5)
```

```{r plot_data2}
smooth_data_fd <- list2funData(smooth_data$smooth)
plot(smooth_data_fd)
```


## Example on simulated data

### On fractional brownian motion

```{r load_brownian, warning=FALSE, message=FALSE}
# Simulate some data
set.seed(42)
fractional_brownian <- generate_fractional_brownian(N = 1000, M = 300, 
                                                    H = 0.5, sigma = 0.05)
```

```{r smooth_data_brown}
smooth_data <- smooth_curves(fractional_brownian, 
                             t0_list = 0.5, 
                             k0_list = 14)
```

```{r estimate_H_brow}
print(smooth_data$parameter$H0)
```


```{r plot_brown}
# Plot a particular observation
plot(fractional_brownian[[1]]$t, fractional_brownian[[1]]$x)
lines(fractional_brownian[[1]]$t, fractional_brownian[[1]]$x_true, col = 'blue', lwd = 2)
lines(smooth_data$smooth[[1]]$t, smooth_data$smooth[[1]]$x, col = 'red', lwd = 2)
```

### On fractional brownian motion

```{r load_piecewise_brownian, warning=FALSE, message=FALSE}
set.seed(42)
piece_frac_brown <- generate_piecewise_fractional_brownian(N = 1000, M = 300, 
                                                           H = c(0.2, 0.5, 0.8), 
                                                           sigma = 0.05)
```

```{r smooth_data_piecebrown}
smooth_data <- smooth_curves(piece_frac_brown, 
                             t0_list = c(0.15, 0.5, 0.85), 
                             k0_list = c(14, 12, 12))
```

```{r estimate H_piecebrow}
print(smooth_data$parameter$H0)
```

```{r plot_piecewise_brow}
# Plot a particular observation
plot(piece_frac_brown[[1]]$t, piece_frac_brown[[1]]$x)
lines(piece_frac_brown[[1]]$t, piece_frac_brown[[1]]$x_true, col = 'blue', lwd = 2)
lines(smooth_data$smooth[[1]]$t, smooth_data$smooth[[1]]$x, col = 'red', lwd = 2)
```